---
title: "Report3"
author: "Salvador Osuna-Caballero"
date: "24/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Describe

Objective: To evaluate if the disease severity to rust is different between ENV and Specie/Material, and if there is interaction between those factors.

```{r, include=FALSE}
# revisa si es necesario instalar los paquetes
list.of.packages <- c("tidyverse",
"ggpubr",
"ggstatsplot",
"openintro",
"rstatix",
"DescTools",
"WRS2",
"metan",
"ggstastplot",
"openxlsx",
"readxl")
new.packages <- list.of.packages[!(list.of.packages %in%
installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
library(tidyverse) # manipulación y visualización de datos
library(ggpubr) # gráficos sencillos
library(ggstatsplot) # gráficos listos para publicar
library(openintro) # datos de ejemplo
library(rstatix) # pruebas P/NP y estadísticos con tuberías.
library(DescTools) # prueba robusta de Yuen
library(WRS2) # tamaño de efecto robustos
library(metan)
library(ggstatsplot)
library(openxlsx)
library(readxl)
```

Solve the alpha lattice design in the field and the RCBD in controlled conditions to create an unique data frame with the BLUPs for disease severity to Rust in Co-18, Co-19, Co-20, Co-all (considering the interaction between ENVs) and Co-CC.

```{r pressure, echo=FALSE}
BLUPs_anova <- read_excel("C:/Users/Salva/Documents/GitHub/Rust-collection/BLUP_ALL_info.xlsx", 
                     sheet = "R", 
                     col_types = 
                       c("text", #GEN
                         "text", #ENV
                         "numeric", #BLUP
                         "text", #Reference
                         "text", #Specie
                         "text", #Name
                         "text", #Origen
                         "text", #Material
                         "text" #Structure
                         ))

BLUPs_anova %>%
group_by(ENV, Material) %>%
get_summary_stats(BLUP, type = "mean_sd")
```

Plotting:

```{r}
ggplot(data = BLUPs_anova, aes(x = Material, y = BLUP)) +
geom_boxplot(aes(fill = ENV), width = 0.8) + theme_classic()
```

```{r}
ggplot(data = BLUPs_anova, aes(x = ENV, y = BLUP)) +
geom_boxplot(aes(fill = Material), width = 0.8) + theme_classic()
```

Normality:

```{r}
ggqqplot(BLUPs_anova, "BLUP", ggtheme = theme_bw()) +
facet_grid(Material ~ ENV)

```

Homogenidad de varianza:

```{r}
BLUPs_anova %>% levene_test(BLUP ~ Material*ENV) 
```

Valores atípicos\

```{r}
BLUPs_anova %>%
  group_by(ENV, Material) %>%
  identify_outliers(BLUP) 
```

Ajustar el modelo

```{r}
BLUPs_anova %>%
  anova_test(BLUP ~ Material)

BLUPs_anova %>%
  anova_test(BLUP ~ Species)

BLUPs_anova %>%
  anova_test(BLUP ~ ENV)
```

Prueba post-hoc para una interacción significativa ( como es el caso)

BLUPs_anova %\>%

pairwise_t\_test(BLUP \~ ENV, p.adjust.method = "holm")%\>%

filter(p.adj \< 0.05)

```{r}
BLUPs_anova %>%
pairwise_t_test(BLUP ~ Material, p.adjust.method = "holm")%>%
filter(p.adj < 0.05)

BLUPs_anova %>%
pairwise_t_test(BLUP ~ ENV, p.adjust.method = "holm")%>%
filter(p.adj < 0.05)

BLUPs_anova %>%
pairwise_wilcox_test(BLUP ~ Species, p.adjust.method = "holm")%>%
filter(p.adj < 0.05)
```

```{r}
BLUPs_anova %>%
group_by(Material) %>%
pairwise_t_test(BLUP ~ ENV, p.adjust.method = "holm") %>%
filter(p.adj < 0.001)
```

```{r}
BLUPs_anova %>%
group_by(Species) %>%
pairwise_t_test(BLUP ~ ENV, p.adjust.method = "holm") %>%
filter(p.adj < 0.001)
```

```{r}
BLUPs_anova %>%
group_by(Species) %>%
pairwise_t_test(ENV ~ BLUP, p.adjust.method = "holm") %>%
filter(p.adj < 0.01)
```

Plotting

```{r}
grouped_ggbetweenstats(data = BLUPs_anova, x = Species, y = BLUP, grouping.var = ENV, type = "np",
results.subtitle = F, messages = F, var.equal = T, p.adjust.method = "holm")
```

```{r}
ggbetweenstats(x = Species, y = BLUP, data = BLUPs_anova, 
               p.adjust.method = "none", type = "np",
               bf.message = FALSE, var.equal = F,
               ggsignif.args = list(textsize = 1.5, tip_length = 0.01)) +
  theme(text = element_text(size = 8), plot.subtitle = element_text(size=8))
```

\
\

\

\
\
